Comparative vaccine effectiveness (ChAdOx1 versus BNT162b2) in Health and Social Care workers

### 2021-09-06 analysis/report/effectiveness_report.Rmd
add covid-19 emergency attendances outcome; remove tests; remove diagnoses

* registered at a GP practice using TPP's SystmOne clinical information system on the day of they received their first dose of BNT162b2 or ChAdOx1; 
* the date of vaccination was between `r format(as.Date(gbl_vars$start_date_az), "%d %B")` and `r format(as.Date(gbl_vars$lastvax_date), "%d %B %Y")` (`r as.Date(gbl_vars$lastvax_date) - as.Date(gbl_vars$start_date_az)` days), a period when both vaccines were being administered widely;
* aged between 18 and 64 inclusive; 
* not classed as Clinically Extremely Vulnerable, as set out by government guidance, at the time of vaccination; 
* information on sex, ethnicity, deprivation, and geographical region was known.

@@

* death or deregistration;
* Second vaccine dose;
* The study end date, `r format(as.Date(gbl_vars$end_date), "%d %B %Y")`.

@@

* SARS-CoV-2 infection, as identified via SGSS records. Both PCR and LFT tests are included.
* Unplanned COVID-19-related hospitalisation via HES records. ICD-10 codes are used to identify COVID-19-related admissions. 
* Unplanned COVID-19-related ICU admission via HES records. ICD-10 codes are used to identify COVID-19-related admissions. 
* COVID-19-related death via death registration records. With COVID-19 ICD-10 codes anywhere on the death certificate (i.e., as an underlying or contributing cause of death)

@@

Patient characteristics used for adjustment include: age, sex, deprivation, ethnicity, NHS region, clinically “at risk” (but not clinically extremely vulnerable) as per national prioritisation guidelines, asthma, number of previous SARS-CoV-2 tests (via SGSS), rurality, evidence of prior covid infection (positive test or COVID-19 hospitalisation), number of recorded comorbidities, severe mental illness. All characteristics are ascertained as at the time of vaccination.

@@

### Time-dependent Cox models
Time-dependent Cox models are used to estimate the time-varying effects for each vaccine type. Time zero is the date of vaccination, i.e., using a treatment time-scale. Outcome events occurring on the same day as vaccination are included in follow-up. Date of vaccination is included as an additional baseline covariate using a restricted cubic spline with knots at the first and second tertile, plus boundary knots. Here, follow-up time is stratified into weekly intervals with the vaccine-specific effect estimated within each time-strata. In `R`'s `coxph` function, this is modelled using `vaccine_type:strata(week_since_first_dose)`.

@@

* *ChAdOx1* versus *BNT162b2* hazard ratio splines
* Risk-adjusted survival curves for *ChAdOx1* and *BNT162b2*

@@

>>>>

-   registered at a GP practice using TPP's SystmOne clinical information system on the day of they received their first dose of BNT162b2 or ChAdOx1;
-   the date of vaccination was between `r format(as.Date(gbl_vars$start_date_az), "%d %B")` and `r format(as.Date(gbl_vars$lastvax_date), "%d %B %Y")` (`r as.Date(gbl_vars$lastvax_date) - as.Date(gbl_vars$start_date_az)` days), a period when both vaccines were being administered widely;
-   aged between 18 and 64 inclusive;
-   not classed as Clinically Extremely Vulnerable, as set out by government guidance, at the time of vaccination;
-   information on sex, ethnicity, deprivation, and geographical region was known.

@@

-   death or deregistration;
-   Second vaccine dose;
-   The study end date, `r format(as.Date(gbl_vars$end_date), "%d %B %Y")`.

@@

-   SARS-CoV-2 infection, as identified via SGSS records. Both PCR and LFT tests are included.
-   Unplanned COVID-19-related hospitalisation via HES records. ICD-10 codes are used to identify COVID-19-related admissions.
-   Unplanned COVID-19-related ICU admission via HES records. ICD-10 codes are used to identify COVID-19-related admissions.
-   COVID-19-related death via death registration records. With COVID-19 ICD-10 codes anywhere on the death certificate (i.e., as an underlying or contributing cause of death)

@@

Patient characteristics used for adjustment include: age, sex, deprivation, ethnicity, NHS region, clinically "at risk" (but not clinically extremely vulnerable) as per national prioritisation guidelines, asthma, number of previous SARS-CoV-2 tests (via SGSS), rurality, evidence of prior covid infection (positive test or COVID-19 hospitalisation), number of recorded comorbidities, severe mental illness. All characteristics are ascertained as at the time of vaccination.

@@

### Time-dependent Cox models

Time-dependent Cox models are used to estimate the time-varying effects for each vaccine type. Time zero is the date of vaccination, i.e., using a treatment time-scale. Outcome events occurring on the same day as vaccination are included in follow-up. Date of vaccination is included as an additional baseline covariate using a restricted cubic spline with knots at the first and second tertile, plus boundary knots. Here, follow-up time is stratified into weekly intervals with the vaccine-specific effect estimated within each time-strata. In `R`'s `coxph` function, this is modelled using `vaccine_type:strata(week_since_first_dose)`.

@@

-   *ChAdOx1* versus *BNT162b2* hazard ratio splines
-   Risk-adjusted survival curves for *ChAdOx1* and *BNT162b2*

@@

cat("  \n\n")
print(glue("### COVID-19 A&E attendance"), "  \n")
wrap_plots(plots_spline("covidemergency"), plots_adjsurv("covidemergency"), nrow=2)
